---
title: "aufgabe 6.5"
author: "David Greussing"
date: "31 10 2020"
output: html_document
---
## Exercise 4: Timeseries Factor Models

Follow the file [tsfmVignette.pdf](https://github.com/braverock/FactorAnalytics/blob/master/vignettes/tsfmVignette.pdf) and interpret your results.

### Theorie
In a time series or macroeconomic factor model, observable economic time series such as industrial production growth rate, interest rates, market returns and inflation are used as common factors that contribute to asset returns. 
- For example, the famous *single index model by Sharpe* (1964) uses the market excess return as the *common factor* (captures economy-wide or market risk) for all assets and the *unexplained returns in the error term* represents the *non-market firm specific risk*. 
- On the other hand, *Chen et al. (1986) uses a multi-factor model* to find that surprise inflation, the spread between long and short-term interest rates and between high and low grade bonds are *significantly priced*, while the market portfolio, aggregate consumption risk and oil price risk are *not priced separately*.

```{r}

install.packages("FactorAnalytics", repos="http://R-Forge.R-project.org")

library(FactorAnalytics)

options(digits=3)

```


```{r}

# The following examples primarily use the managers dataset from the PerformanceAnalytics package. 
# It’s an "xts" data object with:
#                                 - 132 observations of monthly returns
#                                 - on 10 variables:
#                                     - six hypothetical asset managers, 
#                                     - 1x dhec returns (Long-Short Equity hedge fund index)
#                                     - 1x sp500 returns
#                                     - US treasury bonds 10 years (will serve as explanatory factors)
#                                     - US treasury bills 3 months (can be considered as the risk free rate)
#                                 - there are some "not available" observations (start day!)

data(managers)

# We want to see the managers names
colnames(managers)

# and we want to see from when to when the data is available 
range(index(managers))

```

```{r}

# the Ham1-Ham6 are the asset returns we want to explain --> y in our model
asset.names <- colnames(managers[,1:6]) 

# the edhec, sp500 & US Treasury they are the explanatory ones --> x in our model
factor.names <- colnames(managers[,7:9]) 

# Typically, factor models are fit using excess returns. If the asset and factor returns are not in excess return form, "rf.name" can be specified to convert returns into excess returns. 
rf.name <- "US.3m.TR"

# Similarly, market returns can be specified via "mkt.name" to add market-timing factors to the factor model.
mkt.name <- "SP500.TR" 

```

### Let’s take a look at the arguments for *fitTsfm*.

The default model fitting method is *LS regression* and the default variable selection method
is "none" (that is, all factors are included in the model). 
The different model fitting options are: 
- least squares (LS), 
- discounted least squares (DLS) and
- robust regression fitting (Robust)


And variableselection options are:
- "stepwise", 
- "subsets" and 
- "lars"

The default for rf.name and mkt.name are NULL. If rf.name is not specified by the user,
perhaps because the data is already in excess return form, then no risk-free rate adjustment is
made. Similarly, if mkt.name is not specified, market-timing factors are not added to the model.
All other optional control parameters passed through the ellipsis are processed and assimilated
internally by fitTsfm.control.

```{r}

# The series have unequal histories in this sample and “fitTsfm“ removes asset-wise incomplete cases (asset’s return data combined with respective factors’ return data) before fitting a factor model.
args(fitTsfm)

```

```{r}

# Single Index Model using SP500 
fit.singleIndex <- fitTsfm(asset.names=asset.names, 
                           factor.names="SP500.TR",   #specfic factor!
                           rf.name="US.3m.TR", 
                           data=managers)

# Interpretation:
# if the market return rises 1%, then the return of Ham1 rises 0,39%
# R-squared: 1 would be 100% - linear function matches perfectly with the data --> here we have low R-squared

```

```{r}

class(fit.singleIndex)

```

```{r}

names(fit.singleIndex)

```

Overview of the single factor linear fits for the assets. 
```{r}

fit.singleIndex
# Interpretation:
# How good does the single index model fits to the data?
# Ham1 equals a linear regression the most --> fits the best --> R-squared is the highest
# Ham5 does not really fit to this mode --> alfa and R-squared values

```

```{r}

plot(fit.singleIndex, which=12, f.sub=1)

```

### Henriksson-Merton's - market timing models
Market timing accounts for the price movement of the general stock market relative to fixed income securities.
This includes the down.market factor --> max(0, Rf-Rm)
To test market timing ability, this factor can be added to the single index model as shown below. The coefficient of this down-market factor can be
interpreted as the number of "free" put options on the market provided by the manager’s markettimings kills. That is, a negative value for the regression estimate would imply a negative value for market timing ability of the manager.

```{r}

# Henriksson-Merton's market timing model
fit.mktTiming <- fitTsfmMT(asset.names=asset.names, 
                           mkt.name="SP500.TR", # specify which of the columns in data corresponds to the market returns using argument mkt.name.
                           rf.name="US.3m.TR", 
                           data=managers)

t(fit.mktTiming$beta)

# Interpretation:
# when the value of down.market is negative, the ability of market timing of a manager is low --> not even there
# so the manager 2 has the best ability of market timing and after that manager 6 --> they have the hightes intercept (which return they will make when the market makes no return)

```


```{r}

fit.mktTiming$r2
# Interpretation:
# R^2 -> how good the data fits to the model

```

```{r}

fit.mktTiming$resid.sd
# Interpretation:
# volatility: how much it jumps around relative to its relationship to an index(sp500)
# risk: the higher the worse

###fit methods
#ls = least squares
#dls = discounted least squares (weightes least squares)
#robust = is good for data with outliers

```

Fits Model:

The different model fitting options are: 
- (ordinary) least squares (ols / LS) --> Default mode!
- discounted least squares (DLS) and
- robust regression fitting (Robust)

### Ordinary least squares ("ols")

```{r}
#  The next example performs LS regression using all 3 available factors in the dataset.
fit.ols <- fitTsfm(asset.names=asset.names, 
                   factor.names=factor.names, # all 3 available factors: the edhec, sp500 & US.10Y.TR/US Treasury
                   rf.name="US.3m.TR", 
                   data=managers) 


fit.ols$beta
# Interpretation:
# now we consider all factors (explanatory factors)

# Sensitivity: 
# when the return of edhec rises 1% --> Ham2 rises 0,1547%
# when sp500 return rises 1%, Ham2 decreases by 0,195%
# when US.10Y.TR return rises 1%, Ham2 decreases by 0,0504%
```

```{r}

fit.ols$r2
# Interpretation:
# how good does the data fit to the model
# Ham3 fits the best with 66%

```

```{r}

fit.ols$resid.sd
# Interpretation:
# Volatility
# How much they jump around --> most Ham4 0.0427

```


### Discounted least squares ("ols")
DLS is least squares regression using exponentially discounted weights and accounts for time variation in coefficients. Robust regression is resistant to outliers.
```{r}
```

### Other options robust regression ("Robust"). 
```{r}

fit.robust <- fitTsfm(asset.names=asset.names, 
                      factor.names=factor.names, 
                      rf.name="US.3m.TR", 
                      data=managers, 
                      fit.method="Robust") # Method "Robust"!
fit.robust$beta
# Interpretation:

```

```{r}

fit.robust$r2
# Interpretation:
# R-squared is now lower for each
# maybe they all had outliers

```


```{r}

fit.robust$resid.sd
# Interpretation:

```

```{r}

par(mfrow=c(2,1)) 
plot(fit.ols, plot.single=TRUE, which=1, asset.name="HAM3") 
mtext("LS", side=3) 
plot(fit.robust, plot.single=TRUE, which=1, asset.name="HAM3") mtext("Robust", side=3)

# Interpretation:
# volatility is smaller when using the robust fitting method
### variable selection

# lars is a good variable to add
# least angle regression
# it is good when you are afraid of overfitting (that you adjust your model too much)
# when you have high-dimensional data (lots of explanatory factors)

```

```{r}

par(mfrow=c(1,2)) 
plot(fit.ols, which=5, xlim=c(0,0.045), sub="LS") 
plot(fit.robust, which=5, xlim=c(0,0.045), sub="Robust")

```

Though the R-squared values improved by adding more factors in fit.ols (compared to the single index model)

### Variable Selection
One might prefer to employ variable selection methods such as "stepwise", "subsets" or "lars" to avoid over-fitting. The method can be selected via the variable.selection argument. The default "none", uses all the factors and performs no variable selection.
- Specifying *"stepwise"* selects traditional stepwise LS or robust regression using step or step.lmRob respectively. Starting from the given initial set of factors, factors are added (or subtracted) only if the regression fit improves.
- Specifying *"subsets"* enables subsets selection using regsubsets. The best performing subset of any given size or within a range of subset sizes is chosen. Different methods such as exhaustive search (default), forward or backward stepwise, or sequential replacement can be employed.
- Finally, *"lars"* corresponds to least angle regression using lars with variants "lasso" (default), "lar", "forward.stagewise" or "stepwise".

#### LARS = least angle regression 

```{r}

fit.lars <- fitTsfm(asset.names=asset.names, 
                    factor.names=factor.names, 
                    data=managers, 
                    rf.name="US.3m.TR", 
                    variable.selection="lars") 

fit.lars
# Interpretation:
# Subset --> the best performing subset within a range of subset sizes is chosen

```

```{r}

fit.sub <- fitTsfm(asset.names=asset.names, 
                   factor.names=factor.names, 
                   data=managers, 
                   rf.name="US.3m.TR", 
                   variable.selection="subsets", 
                   nvmin=2, nvmax=2) 
# Here, the best subset of size 2 for each asset is chosen by specifying nvmin = nvmax = 2. Note that when nvmin < nvmax, the best subset is chosen from a range of subset sizes [nvmin, nvmax]. Default is nvmin = 1.

# Interpretation:
# we see all together
# intercepts = alpha
# where we see the indices --> betas

```

```{r}

plot(fit.sub, which=2, f.sub=1:3)

```

```{r}

plot(fit.lars, which=2, f.sub=1:3)

```

Comparing the *coefficients* and *R-squared values* from the two models, we find that the method that uses *more factors* for an asset have higher R-squared values as expected. However, when both "lars" and "subsets" chose the same number of factors, "lars" fits have a slightly higher R-squared values.


###  S3 generic methods
Many useful generic accessor functions are available for "tsfm" fit objects:
- coef() returns a matrix of estimated model coefficients including the intercept. 
- fitted() returns an xts data object of the component of asset returns explained by the factor model. 
- residuals() returns an xts data object with the component of asset returns not explained by the factor model. 
- predict() uses the fitted factor model to estimate asset returns given a set of new or simulated factor return data.
- summary() prints standard errors and t-statistics for all estimated coefficients in addition to R-squared values and residual volatilities. 

Argument se.type, one of "Default", "HC" or "HAC", allows for heteroskedasticity and auto-correlation consistent estimates and standard errors whenever possible. A "summary.tsfm" object is returned which contains a list of summary objects returned by "lm", "lm.Rob" or "lars" for each asset fit.

```{r}

methods(class="tsfm")

```

All estimated coefficients from the LS fit using all 3 factors
```{r}

coef(fit.ols)

```

Compare returns data with fitted and residual values for HAM1 from fit.lars

```{r}

HAM1.ts <- merge(fit.lars$data[,1], 
                 fitted(fit.lars)[,1], 
                 residuals(fit.lars)[,1]) 

colnames(HAM1.ts) <- c("HAM1.return","HAM1.fitted","HAM1.residual") 

tail(HAM1.ts)

# Interpretation:
# fitted --> the returns which can be explained through the model
# residual --> the returns which cannot be explained through the model
```

### Summary for fit.sub computing HAC standard erros

```{r}

summary(fit.sub, se.type="HAC")

```
